\vspace{-.6in}Annotation-Based Empirical Performance Tuning\\ Using Orio

### 2009-02-12 doc/papers/ipdps09/results.tex
git-svn-id: https://svn.mcs.anl.gov/repos/performance/orio@428 e7ad4b5f-b827-0410-872f-f7f4bc3d1efb

\begin{figure}[thb] 
\begin{center} 
  \subfigure[Sequential (single core)]{ 
  \includegraphics[width=.4\textwidth]{figures/axpy4_bgp/seq.eps}  
  \label{fig:axpy4-bgp-seq} 
  } 
  \subfigure[Parallel (four cores)]{ 
  \includegraphics[width=.4\textwidth]{figures/axpy4_bgp/par.eps}  
  \label{fig:axpy4-bgp-par} 
  } 
\end{center}
\vspace{-.2in} 
\caption{Performance of AXPY-4 operations on the Blue Gene/P.} 
\label{fig:axpy4-bgp-results} 
\end{figure}

@@

@@

The SpMV operation computes $\forall_{A_{i,j}} \neq 0:y_{i}
\leftarrow y_{i} + A_{i,j} \cdot x_{j}$, where $A$ is a sparse matrix,
and $x$, $y$ are dense vectors. Each element of $A$ is used precisely once
and element reuse is only possible for $x$ and $y$. Thus, to optimize SpMV,
one should use compact data structures for $A$ and try to maximize temporal
reuse of $x$ and $y$. One of the most commonly used data structures for
storing a sparse matrix is compressed sparse row (CSR) storage,
%~\cite{vuduc-thesis} 
illustrated in Figure~\ref{fig:spmv}(a). Elements in each row of $A$ are
packed together in a dense array, \texttt{Aval}, and a corresponding array of
integers \texttt{Aind} stores the column indices. The \texttt{Aptr} array
designates where each sparse row begins in \texttt{Aval} and
\texttt{Aind}. An implementation of SpMV using CSR storage is
shown in Figure~\ref{fig:spmv}(b).

\begin{figure}[tb]
\centering
\vspace{-.2in}
\begin{tabular}{cc}
\begin{minipage}[b]{.5\textwidth}
\centering
\includegraphics[width=.8\textwidth]{figures/spmv.eps}  
\end{minipage}
&
\begin{minipage}[b]{.5\textwidth}
\scriptsize
\begin{verbatim}
for (i=0; i<num_rows; i++)
  for (j=Aptr[i]; j<Aptr[i+1]; j++)
    y[i] += Aval[j]*x[Aind[j]];

\end{verbatim}
\end{minipage}\\
(a) & (b)\\
\end{tabular}
\vspace{-.2in}
\caption{(a) Compressed sparse row (CSR) format (b) Basic implementation of CSR-based SpMV.}
\label{fig:spmv}
%\vspace{.2in}
\end{figure}

@@

Figure~\ref{fig:ex27-cookie-results} shows the performance results on the
Intel Xeon. The problem size is the number of grid points in the $x$ and $y$
directions, not the size of the sparse input matrix. The matrix dimension is
based the grid size and the number of field components computed for each grid
point, for example, an $8 \times 8$ problem involves sparse matrices of
dimension 900 with 17,040 nonzero entries. In this experiment, we ran in
three different multicore modes: SMP (one MPI process, eight
threads/process), dual (four MPI processes, two threads/process), and virtual
node or VN (eight MPI processes, one thread/process). The code tuned by Orio
consistently outperforms the alternatives. Furthermore, the performance of
the hand-tuned code is almost equivalent to that of the icc-optimized simple
code. Performance differences between the Orio version and the hand-tuned
version become more significant as more threads per process are employed to
facilitate OpenMP parallelization. The SMP and VN results show that
optimizations using loop-level parallelism (through OpenMP) achieve much
better performance than using MPI coarse-grained parallelism.
%Moreover, for the multithread cases, tuning using Orio determined that
%the performance benefit of SIMDization is more significant than that of
%thread-level parallelization only when the problem sizes are small. The
%reason for this is due to the overhead introduced by OpenMP from coordinating
%its threads across processors.
%So, each optimized
%code for the many-thread cases comprises two distinct tuned codes to handle
\begin{figure} [ht]
\begin{center} 
  \subfigure[SMP: $p=1$, $t/p=8$]{
  \includegraphics[width=.3\textwidth]{figures/ex27_cookie/smp.eps}
  \label{fig:ex27-cookie-smp} } \subfigure[Dual: $p=4$, $t/p=2$]{
  \includegraphics[width=.3\textwidth]{figures/ex27_cookie/dual.eps}
  \label{fig:ex27-cookie-dual} } \subfigure[VN: $p=8$, $t/p=1$]{
  \includegraphics[width=.3\textwidth]{figures/ex27_cookie/vn.eps}
  \label{fig:ex27-cookie-vn} }
\end{center}
\vspace{-.2in} 
\caption{Performance of inode SpMV on eight-core Intel machine; $p$ is the number of processes, and $t/p$ is the threads per process.} 
\label{fig:ex27-cookie-results} 
\end{figure}

@@

\begin{figure} [htb]
\begin{center} 
  \subfigure[1 node, SMP mode]{ 
  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/n1_smp.eps}  
  \label{fig:ex27-bgp-smp-n1} 
  } 
  \subfigure[1 node, Dual mode]{ 
  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/n1_dual.eps}  
  \label{fig:ex27-bgp-dual-n1} 
  } 
  \subfigure[1 node, VN mode]{ 
  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/n1_vn.eps}  
  \label{fig:ex27-bgp-vn-n1} 
  } 
  \subfigure[8 nodes, SMP mode]{ 
  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/n8_smp.eps}  
  \label{fig:ex27-bgp-smp-n8} 
  } 
  \subfigure[8 nodes, Dual mode]{ 
  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/n8_dual.eps}  
  \label{fig:ex27-bgp-dual-n8} 
  } 
  \subfigure[8 nodes, VN mode]{ 
  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/n8_vn.eps}  
  \label{fig:ex27-bgp-vn-n8} 
  } 
%  \subfigure[32 nodes, SMP mode]{ 
%  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/n32_smp.eps}  
%  \label{fig:ex27-bgp-smp-n32} 
%  } 
%  \subfigure[32 nodes, Dual mode]{ 
%  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/n32_dual.eps}  
%  \label{fig:ex27-bgp-dual-n32} 
%  } 
%  \subfigure[32 nodes, VN mode]{ 
%  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/n32_vn.eps}  
%  \label{fig:ex27-bgp-vn-n32} 
%  } 
\end{center}
\vspace{-.2in} 
\caption{Performance of inode SpMV on Blue Gene/P for 1, 8, and 32 nodes.} 
\label{fig:ex27-bgp-results1} 
\end{figure}

\comment{
\begin{figure} 
\begin{center} 
  \subfigure[32x32, SMP mode]{ 
  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/s32_smp.eps}  
  \label{fig:ex27-bgp-smp-32x32} 
  } 
  \subfigure[32x32, Dual mode]{ 
  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/s32_dual.eps}  
  \label{fig:ex27-bgp-dual-32x32} 
  } 
  \subfigure[32x32, VN mode]{ 
  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/s32_vn.eps}  
  \label{fig:ex27-bgp-vn-32x32} 
  } 
  \subfigure[64x64, SMP mode]{ 
  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/s64_smp.eps}  
  \label{fig:ex27-bgp-smp-64x64} 
  } 
  \subfigure[64x64, Dual mode]{ 
  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/s64_dual.eps}  
  \label{fig:ex27-bgp-dual-64x64} 
  } 
  \subfigure[64x64, VN mode]{ 
  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/s64_vn.eps}  
  \label{fig:ex27-bgp-vn-64x64} 
  } 
  \subfigure[128x128, SMP mode]{ 
  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/s128_smp.eps}  
  \label{fig:ex27-bgp-smp-128x128} 
  } 
  \subfigure[128x128, Dual mode]{ 
  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/s128_dual.eps}  
  \label{fig:ex27-bgp-dual-128x128} 
  } 
  \subfigure[128x128, VN mode]{ 
  \includegraphics[width=.3\textwidth]{figures/ex27_bgp/s128_vn.eps}  
  \label{fig:ex27-bgp-vn-128x128} 
  } 
\end{center}
\vspace{-.2in}
\caption{Performance of inode SpMV on Blue Gene/P for problem sizes 32x32, 64x64, and 128x128.} 
\label{fig:ex27-bgp-results2} 
\end{figure} 
}

@@

\subsubsection{2-D Finite-Difference Time-Domain Method for Computational Electromagnetics}  
We consider the two-dimensional Finite Difference Time Domain (FDTD) algorithm, a
popular method for solving the time-dependent Maxwell's equations in the
context of computational 
%
\begin{wrapfigure}{r}{3in}
\vspace{-0.2in}
\begin{center}
\begin{minipage}{3in} 
\scriptsize
\begin{verbatim} 
for(t=0; t<tmax; t++) { 
  for (j=0; j<ny; j++) ey[0][j] = t; 
  for (i=1; i<nx; i++) 
    for (j=0; j<ny; j++) 
      ey[i][j] -= 0.5*(hz[i][j] - hz[i-1][j]); 
  for (i=0; i<nx; i++) 
    for (j=1; j<ny; j++) 
      ex[i][j] -= 0.5*(hz[i][j] - hz[i][j-1]); 
  for (i=0; i<nx; i++) 
    for (j=0; j<ny; j++) 
      hz[i][j] -= 0.7*(ex[i][j+1] - ex[i][j]
                  + ey[i+1][j] - ey[i][j]); 
} 
\end{verbatim} 
\end{minipage} 
\end{center}
\vspace{-0.2in}
\caption{2-D FDTD code.} 
\label{fig:fdtd-2d-code} 
\vspace{-.1in}
\end{wrapfigure}
%
electrodynamic problems. As shown in Figure~\ref{fig:fdtd-2d-code}, the 2-D
FDTD method is implemented as an outer iteration over time containing four imperfectly
nested loops.  The arrays $ex$ and $ey$ denote the electric field components,
and the array $hz$ denotes the magnetic field.

@@

\begin{figure}[htb] 
\begin{center} 
  \subfigure[Sequential (T=500)]{ 
  \includegraphics[width=.4\textwidth]{figures/fdtd-2d-cookie/seq.eps} 
  \label{fig:fdtd-2d-cookie-seq} } \subfigure[Parallel (T=500, 
  N=2000)]{ 
  \includegraphics[width=.4\textwidth]{figures/fdtd-2d-cookie/par.eps} 
  \label{fig:fdtd-2d-cookie-par} } 
\end{center}
\vspace{-.2in} 
\caption{2-D FDTD performance on an eight-core Intel Xeon.} 
\label{fig:fdtd-2d-cookie-results} 
%\vspace{-.2in}
\end{figure}

@@

\subsubsection{3-D Gauss-Seidel Successive Over-Relaxation Method}  
Figure~\ref{fig:seidel-code} shows the 3-D Gauss-Seidel computation, 
which is sometimes is referred to as
\emph{successive displacement method}, indicating the dependence of the 
iterations on the ordering. If the ordering is
%
%\begin{figure}[ht]
\begin{wrapfigure}{r}{3in}
\vspace{-0.3in}
\begin{center}
\begin{minipage}{3in} 
\scriptsize
\begin{verbatim} 
for (t=0; t<=T-1; t++) 
  for (i=1; i<=N-2; i++) 
    for (j=1; j<=N-2; j++) 
      A[i][j] = (A[i-1][j-1] + A[i-1][j] 
          + A[i-1][j+1] + A[i][j-1] + A[i][j] 
          + A[i][j+1] + A[i+1][j-1] + A[i+1][j] 
          + A[i+1][j+1]) / 9.0; 
\end{verbatim} 
\end{minipage} 
\end{center}
\vspace{-0.2in}
\caption{3-D Gauss-Seidel code} 
\label{fig:seidel-code} 
\vspace{-.1in}
\end{wrapfigure}
%
changed, the components of the new iterations will change as well.

@@

\begin{figure} [htb]
\begin{center} 
  \subfigure[Sequential (T=500)]{ 
  \includegraphics[width=.4\textwidth]{figures/seidel-cookie/seq.eps} 
  \label{fig:seidel-cookie-seq} } \subfigure[Parallel (T=500, 
  N=4000)]{ 
  \includegraphics[width=.4\textwidth]{figures/seidel-cookie/par.eps} 
  \label{fig:seidel-cookie-par} } 
\end{center}
\vspace{-.2in} 
\caption{3-D Gauss-Seidel performance on eight-core Intel machine.} 
\label{fig:seidel-cookie-results} 
\vspace{-.1in}
\end{figure}

@@

LU factorization or decomposition is a numerical method for the solution of linear
systems 
%
\begin{wrapfigure}{r}{3in}
\begin{center}
\vspace{-.2in}
\begin{minipage}{2.8in} 
\scriptsize
\begin{verbatim} 
for (k=0; k<=N-1; k++) { 
  for (j=k+1; j<=N-1; j++) 
    A[k][j] /= A[k][k]; 
  for(i=k+1; i<=N-1; i++) 
    for (j=k+1; j<=N-1; j++) 
      A[i][j] -= A[i][k]*A[k][j]; 
} 
\end{verbatim} 
\end{minipage} 
\end{center}
\vspace{-0.2in}
\caption{LU Decomposition code.} 
\label{fig:lu-code} 
\end{wrapfigure}
%
of equations; a simple implementation is shown in Figure~\ref{fig:lu-code}.

@@

\begin{figure}[htb]
\begin{center} 
  \subfigure[Sequential]{ 
  \includegraphics[width=.4\textwidth]{figures/lu-cookie/seq.eps} 
  \label{fig:lu-cookie-seq} } \subfigure[Parallel (N=4000)]{ 
  \includegraphics[width=.4\textwidth]{figures/lu-cookie/par.eps} 
  \label{fig:lu-cookie-par} } 
\end{center} 
\vspace{-.2in}
\caption{LU Decomposition performance on eight-core Intel machine.} 
\label{fig:lu-cookie-results} 
\end{figure}

>>>>

\begin{figure*}%[t] 
\begin{center} 
  \subfigure[Sequential (single core)]{ 
  \includegraphics[width=.45\textwidth]{figures/axpy4_bgp/seq.eps}  
  \label{fig:axpy4-bgp-seq} 
  } 
  \subfigure[Parallel (four cores)]{ 
  \includegraphics[width=.45\textwidth]{figures/axpy4_bgp/par.eps}  
  \label{fig:axpy4-bgp-par} 
  } 
\end{center}
\caption{Performance of AXPY-4 operations on the Blue Gene/P.} 
\label{fig:axpy4-bgp-results} 
\end{figure*}

@@

\begin{figure}%[tb]
\begin{tabular}{c}
\begin{minipage}[b]{.47\textwidth}
\centering
\includegraphics[width=1.0\textwidth]{figures/spmv.eps}  
\end{minipage}
\\
{\footnotesize (a)}
\\
\begin{minipage}[b]{.3\textwidth}
\scriptsize
\begin{verbatim}

for (i=0; i<num_rows; i++)
  for (j=Aptr[i]; j<Aptr[i+1]; j++)
    y[i] += Aval[j]*x[Aind[j]];

\end{verbatim}
\end{minipage}
\\
{\footnotesize (b)}
\\
\end{tabular}

\caption{(a) Compressed sparse row (CSR) format (b) Basic implementation of CSR-based SpMV.}
\label{fig:spmv}
\end{figure}

@@

The SpMV operation computes $\forall_{A_{i,j}} \neq 0:y_{i}
\leftarrow y_{i} + A_{i,j} \cdot x_{j}$, where $A$ is a sparse matrix,
and $x$, $y$ are dense vectors. Each element of $A$ is used precisely once
and element reuse is only possible for $x$ and $y$. Thus, to optimize SpMV,
one should use compact data structures for $A$ and try to maximize temporal
reuse of $x$ and $y$. One of the most commonly used data structures for
storing a sparse matrix is compressed sparse row (CSR) storage
~\cite{vuduc-thesis},
illustrated in Figure~\ref{fig:spmv}(a). Elements in each row of $A$ are
packed together in a dense array, \texttt{Aval}, and a corresponding array of
integers \texttt{Aind} stores the column indices. The \texttt{Aptr} array
designates where each sparse row begins in \texttt{Aval} and
\texttt{Aind}. An implementation of SpMV using CSR storage is
shown in Figure~\ref{fig:spmv}(b).

@@

Figure~\ref{fig:ex27-cookie-results} shows the performance results on the
Intel Xeon. The problem size is the number of grid points in the $x$ and $y$
directions, not the size of the sparse input matrix. The matrix dimension is
based the grid size and the number of field components computed for each grid
point, for example, an $8 \times 8$ problem involves sparse matrices of
dimension 900 with 17,040 nonzero entries. In this experiment, we ran in
three different multicore modes: SMP (one MPI process, eight
threads/process), dual (four MPI processes, two threads/process), and virtual
node or VN (eight MPI processes, one thread/process). The code tuned by Orio
consistently outperforms the alternatives. Furthermore, the performance of
the hand-tuned code is almost equivalent to that of the icc-optimized simple
code. Performance differences between the Orio version and the hand-tuned
version become more significant as more threads per process are employed to
facilitate OpenMP parallelization. The SMP and VN results show that
optimizations using loop-level parallelism (through OpenMP) achieve much
better performance than using MPI coarse-grained parallelism.
%Moreover, for the multithread cases, tuning using Orio determined that
%the performance benefit of SIMDization is more significant than that of
%thread-level parallelization only when the problem sizes are small. The
%reason for this is due to the overhead introduced by OpenMP from coordinating
%its threads across processors.
%So, each optimized
%code for the many-thread cases comprises two distinct tuned codes to handle

\begin{figure*}%[ht]
\begin{center} 
  \subfigure[SMP: $p=1$, $t/p=8$]{
  \includegraphics[width=.315\textwidth]{figures/ex27_cookie/smp.eps}
  \label{fig:ex27-cookie-smp} } \subfigure[Dual: $p=4$, $t/p=2$]{
  \includegraphics[width=.315\textwidth]{figures/ex27_cookie/dual.eps}
  \label{fig:ex27-cookie-dual} } \subfigure[VN: $p=8$, $t/p=1$]{
  \includegraphics[width=.315\textwidth]{figures/ex27_cookie/vn.eps}
  \label{fig:ex27-cookie-vn} }
\end{center}
\caption{Performance of inode SpMV on eight-core Intel machine; $p$ is the number of processes, and $t/p$ is the threads per process.} 
\label{fig:ex27-cookie-results} 
\end{figure*}

@@

\begin{figure*} %[htb]
\begin{center} 
  \subfigure[1 node, SMP mode]{ 
  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/n1_smp.eps}  
  \label{fig:ex27-bgp-smp-n1} 
  } 
  \subfigure[1 node, Dual mode]{ 
  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/n1_dual.eps}  
  \label{fig:ex27-bgp-dual-n1} 
  } 
  \subfigure[1 node, VN mode]{ 
  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/n1_vn.eps}  
  \label{fig:ex27-bgp-vn-n1} 
  } 
  \subfigure[8 nodes, SMP mode]{ 
  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/n8_smp.eps}  
  \label{fig:ex27-bgp-smp-n8} 
  } 
  \subfigure[8 nodes, Dual mode]{ 
  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/n8_dual.eps}  
  \label{fig:ex27-bgp-dual-n8} 
  } 
  \subfigure[8 nodes, VN mode]{ 
  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/n8_vn.eps}  
  \label{fig:ex27-bgp-vn-n8} 
  } 
%  \subfigure[32 nodes, SMP mode]{ 
%  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/n32_smp.eps}  
%  \label{fig:ex27-bgp-smp-n32} 
%  } 
%  \subfigure[32 nodes, Dual mode]{ 
%  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/n32_dual.eps}  
%  \label{fig:ex27-bgp-dual-n32} 
%  } 
%  \subfigure[32 nodes, VN mode]{ 
%  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/n32_vn.eps}  
%  \label{fig:ex27-bgp-vn-n32} 
%  } 
\end{center}
\caption{Performance of inode SpMV on Blue Gene/P for 1 and 8 nodes.} 
\label{fig:ex27-bgp-results1} 
\end{figure*}

\comment{
\begin{figure} 
\begin{center} 
  \subfigure[32x32, SMP mode]{ 
  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/s32_smp.eps}  
  \label{fig:ex27-bgp-smp-32x32} 
  } 
  \subfigure[32x32, Dual mode]{ 
  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/s32_dual.eps}  
  \label{fig:ex27-bgp-dual-32x32} 
  } 
  \subfigure[32x32, VN mode]{ 
  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/s32_vn.eps}  
  \label{fig:ex27-bgp-vn-32x32} 
  } 
  \subfigure[64x64, SMP mode]{ 
  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/s64_smp.eps}  
  \label{fig:ex27-bgp-smp-64x64} 
  } 
  \subfigure[64x64, Dual mode]{ 
  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/s64_dual.eps}  
  \label{fig:ex27-bgp-dual-64x64} 
  } 
  \subfigure[64x64, VN mode]{ 
  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/s64_vn.eps}  
  \label{fig:ex27-bgp-vn-64x64} 
  } 
  \subfigure[128x128, SMP mode]{ 
  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/s128_smp.eps}  
  \label{fig:ex27-bgp-smp-128x128} 
  } 
  \subfigure[128x128, Dual mode]{ 
  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/s128_dual.eps}  
  \label{fig:ex27-bgp-dual-128x128} 
  } 
  \subfigure[128x128, VN mode]{ 
  \includegraphics[width=.315\textwidth]{figures/ex27_bgp/s128_vn.eps}  
  \label{fig:ex27-bgp-vn-128x128} 
  } 
\end{center}
\caption{Performance of inode SpMV on Blue Gene/P for problem sizes 32x32, 64x64, and 128x128.} 
\label{fig:ex27-bgp-results2} 
\end{figure} 
}

@@

\subsubsection{2-D Finite-Difference Time-Domain Method for Computational Electromagnetics}  
We consider the two-dimensional Finite Difference Time Domain (FDTD)
algorithm, a popular method for solving the time-dependent Maxwell's
equations in the context of computational electrodynamic problems. As
shown in Figure~\ref{fig:fdtd-2d-code}, the 2-D FDTD method is
implemented as an outer iteration over time containing four
imperfectly nested loops.  The arrays $ex$ and $ey$ denote the
electric field components, and the array $hz$ denotes the magnetic
field.

%
\begin{figure}
\begin{center}
\begin{minipage}{3in} 
\scriptsize
\begin{verbatim} 
for(t=0; t<tmax; t++) { 
  for (j=0; j<ny; j++) ey[0][j] = t; 
  for (i=1; i<nx; i++) 
    for (j=0; j<ny; j++) 
      ey[i][j] -= 0.5*(hz[i][j] - hz[i-1][j]); 
  for (i=0; i<nx; i++) 
    for (j=1; j<ny; j++) 
      ex[i][j] -= 0.5*(hz[i][j] - hz[i][j-1]); 
  for (i=0; i<nx; i++) 
    for (j=0; j<ny; j++) 
      hz[i][j] -= 0.7*(ex[i][j+1] - ex[i][j]
                  + ey[i+1][j] - ey[i][j]); 
} 
\end{verbatim} 
\end{minipage} 
\end{center}
\caption{2-D FDTD code.} 
\label{fig:fdtd-2d-code} 
\end{figure}
%

@@

\begin{figure*}%[htb] 
\begin{center} 
  \subfigure[Sequential (T=500)]{ 
  \includegraphics[width=.45\textwidth]{figures/fdtd-2d-cookie/seq.eps} 
  \label{fig:fdtd-2d-cookie-seq} } \subfigure[Parallel (T=500, 
  N=2000)]{ 
  \includegraphics[width=.45\textwidth]{figures/fdtd-2d-cookie/par.eps} 
  \label{fig:fdtd-2d-cookie-par} } 
\end{center}
\caption{2-D FDTD performance on an eight-core Intel Xeon.} 
\label{fig:fdtd-2d-cookie-results} 
\end{figure*}

\begin{figure*}%[htb]
\begin{center} 
  \subfigure[Sequential (T=500)]{ 
  \includegraphics[width=.45\textwidth]{figures/seidel-cookie/seq.eps} 
  \label{fig:seidel-cookie-seq} } \subfigure[Parallel (T=500, 
  N=4000)]{ 
  \includegraphics[width=.45\textwidth]{figures/seidel-cookie/par.eps} 
  \label{fig:seidel-cookie-par} } 
\end{center}
\caption{3-D Gauss-Seidel performance on eight-core Intel machine.} 
\label{fig:seidel-cookie-results} 
\end{figure*}

\begin{figure*}%[htb]
\begin{center} 
  \subfigure[Sequential]{ 
  \includegraphics[width=.45\textwidth]{figures/lu-cookie/seq.eps} 
  \label{fig:lu-cookie-seq} } \subfigure[Parallel (N=4000)]{ 
  \includegraphics[width=.45\textwidth]{figures/lu-cookie/par.eps} 
  \label{fig:lu-cookie-par} } 
\end{center} 
\caption{LU Decomposition performance on eight-core Intel machine.} 
\label{fig:lu-cookie-results} 
\end{figure*}

@@

\subsubsection{3-D Gauss-Seidel Successive Over-Relaxation Method}  
Figure~\ref{fig:seidel-code} shows the 3-D Gauss-Seidel computation, 
which is sometimes is referred to as
\emph{successive displacement method}, indicating the dependence of the 
iterations on the ordering. If the ordering is changed, the components
of the new iterations will change as well.
%
%\begin{figure}[ht]
\begin{figure}%[ht]
\begin{center}
\begin{minipage}{3in} 
\scriptsize
\begin{verbatim} 
for (t=0; t<=T-1; t++) 
  for (i=1; i<=N-2; i++) 
    for (j=1; j<=N-2; j++) 
      A[i][j] = (A[i-1][j-1] + A[i-1][j] 
          + A[i-1][j+1] + A[i][j-1] + A[i][j] 
          + A[i][j+1] + A[i+1][j-1] + A[i+1][j] 
          + A[i+1][j+1]) / 9.0; 
\end{verbatim} 
\end{minipage} 
\end{center}
\caption{3-D Gauss-Seidel code} 
\label{fig:seidel-code} 
\end{figure}
%

@@

@@

LU factorization or decomposition is a numerical method for the
solution of linear systems of equations; a simple implementation is
shown in Figure~\ref{fig:lu-code}.
%
\begin{figure}%[ht]
\begin{center}
\begin{minipage}{2.8in} 
\scriptsize
\begin{verbatim} 
for (k=0; k<=N-1; k++) { 
  for (j=k+1; j<=N-1; j++) 
    A[k][j] /= A[k][k]; 
  for(i=k+1; i<=N-1; i++) 
    for (j=k+1; j<=N-1; j++) 
      A[i][j] -= A[i][k]*A[k][j]; 
} 
\end{verbatim} 
\end{minipage} 
\end{center}
\caption{LU Decomposition code.} 
\label{fig:lu-code} 
\end{figure}
%

@@
